Wildcard: Spreadsheet-Driven Customization of Web Applications

### 2020-01-14 paper.tex
done

Web applications often don't meet the precise needs of their users.
Sometimes there is a browser extension available to patch an issue, and
if the user is a programmer they might be able to fix it themselves. But
for most people, the only recourse is to complain to the developers, or
more likely, to simply give up. Back in 1977, in \emph{Personal Dynamic
Media} \autocite{kay1977}, Alan Kay envisioned personal computing as a
medium that let a user ``mold and channel its power to his own needs,''
but today, software behaves more like concrete than clay.

@@

We have implemented this technique in a prototype browser extension
called Wildcard, and used it to build demos which suggest that this
paradigm can support useful customizations, ranging from sorting lists
of data to adding whole new features to applications (shown in
Section~\ref{sec:demos}).

@@

In addition to fetching data from other sources, Wildcard formulas can
also perform computations. In this example, the user wants to augment
the TodoMVC todo list app with a ``snooze'' feature, which will
temporarily hide a todo from the list until a certain date.{
Figure~\ref{fig:todomvc-demo} shows an overview of this customization.}

@@

In this section we've presented just a few use cases for
spreadsheet-driven customization, to suggest some possibilities of the
paradigm. We think the spreadsheet model is flexible enough to support a
wide range of other useful modifications, while remaining familiar and
easy to use.

@@

  \CommentTok{// Function to extract DOM element:}
\NormalTok{  el}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \VariableTok{row}\OperatorTok{.}\FunctionTok{querySelector}\NormalTok{(}\VerbatimStringTok{\textasciigrave{}.\_1ebt2xej\textasciigrave{}}\NormalTok{)}\OperatorTok{,}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

Sometimes, more advanced scraping techniques are necessary to extract
the necessary data from the application. For example, we have prototyped
a mechanism for site adapters to observe AJAX requests made by the
browser and extract data directly from JSON responses. This mechanism
was used to implement the Airbnb Walkscore example, since latitude and
longitude aren't shown in the Airbnb UI, but they are available in AJAX
responses. This technique seems promising because AJAX responses already
contain data in a structured form, and web applications are increasingly
loading data using AJAX. Another technique we might consider is enabling
site adapters to scrape data across multiple pages for paginated lists
of results (as explored in \autocite{huynh2006}).

@@

First, the site adapter needs to support sending updates from the table
to the original page. We have made this fairly straightforward, because
most DOM manipulation is not performed directly by the site
adapter---instead, Wildcard automatically mutates the DOM to reflect the
spreadsheet state, using the same declarative spec shown above. The only
exception to this is row actions (like favoriting an Airbnb listing),
which are implemented as Javascript functions that can perform arbitrary
behaviors like simulating clicks on buttons or mutating the DOM.

Another challenge is triggering updates to the spreadsheet data in
response to UI changes that happen after initial page load. Site
adapters are responsible for implementing site-specific logic for
recognizing these changes by observing the DOM. So far, we have been
able to use event listeners and the MutationObserver API to successfully
observe changes, but it may prove challenging to correctly respond to
updates on some sites purely by observing the DOM.

@@

The idea of spreadsheet-driven customization is guided by several design
principles, inspired by prior work and our own experimentation. We think
these principles might also broadly inform the design of tools for end
user software customization.

@@

Computing does not need to be organized this way. For example, UNIX
offers a compelling alternative design: many small single-purpose
programs, all of which manipulate a universal format of text streams.
Users get high leverage from mastering a text editor and other text
manipulation tools, since they can be applied to a huge variety of
tasks. A user's preferred text editor can even serve as an interactive
input mechanism in shell programs, e.g.~for editing git commit messages.

Spreadsheet-driven customization aims to bring some of the UNIX ethos to
the world of isolated Web applications, by creating a consistent data
structure to represent the data inside many applications. In UNIX, the
universal format is a text stream; in Wildcard, it is a relational
table. Because Wildcard maps the data from all applications to the table
format, users can master the tools provided by Wildcard and apply them
to customize many different applications.

This idea relates to Beaudouin-Lafon and Mackay's notion of
\emph{polymorphic interface instruments} \autocite{beaudouin-lafon2000},
which are UI elements that can be used in different contexts (for
example, a color picker that can be used in many different drawing
applications). It also relates to ideas of literacy in a medium. DiSessa
notes that textual literacy depends on the fact that writing can be
adapted to many different genres and uses \autocite{disessa2000}; if
people needed to relearn reading and writing from scratch when switching
from essays to emails, the medium would lose its potency. Having generic
tools is especially important for software customization, because a
common barrier to customizing software is not having enough time
\autocite{mackay1991}. It's more likely that people will customize
software regularly if they can reuse the same tools across many
applications.

This design principle leads to several challenges. First, any universal
abstraction has its constraints, and may not naturally express the data
in every application. We plan to explore the limits of the table
abstraction further, by trying to build adapters for more sites with
varied data formats. We expect that many types of data will fit easily
into tables: lists of search results, news articles, and messages can
all naturally be seen as relations. Documents (e.g.~Google Docs) or
graphs (e.g.~social network friend graphs) may prove more challenging to
map to the table abstraction.

@@

One of the most interesting properties of spreadsheets is that users
familiar with only a tiny sliver of their functionality (e.g., storing
tables of numbers and computing simple sums) can still use them in
genuinely valuable ways. This supports the user's natural motivation to
continue using the tool, and to eventually learn its more powerful
features if needed \autocite{nardi1991}. In contrast, many traditional
programming systems require an enormous upfront time investment before
someone is able to write a program that helps them achieve a useful
task.

@@

Another aspect of providing a low floor is providing an ``in-place
toolchain'' \autocite{2019}: minimizing the effort of moving from using
to customizing, by making customization tools available in the same
environment where the user is already using the software. This quality
is distinct from the level of technical skill needed to use the tool.
For example, setting up a workflow in an end user programming system
like \href{https://ifttt.com/}{IFTTT} does not require much programming
skill, but it does require leaving the user's normal software and
entering a separate environment; conversely, running a Javascript
snippet in the browser console requires programming skills, but can be
done immediately in the flow of using a website.

@@

Since we have still only built a few site adapters and demos, it is
still too early to tell how high the ceiling is for the customizations
that can be achieved with Wildcard. But we think that with enough
built-in formulas (both standard spreadsheet formulas for computation,
and formulas for fetching data from other sources), the language could
support a wide variety of customizations. We plan to explore this aspect
further by trying to solve real problems and observing where limitations
emerge in practice. Section~\ref{sec:demos} also discusses some
limitations we expect to face.

@@

Real-world spreadsheet usage is highly collaborative. Many users just
perform simple changes, while their coworkers help with writing more
complex formulas or even programming macros \autocite{nardi1990}.
Inspired by this, we aim to make spreadsheet-driven customization a
collaborative activity that combines the different abilities of
different users.

@@

The first party developers of the original software can also play a role
in enabling customization. Spreadsheet-driven customization does not
depend on cooperation from first-party website developers, but if they
were to expose structured data in their web application clients, it
would eliminate the need for third party site adapters. We think there
are compelling reasons for first parties to consider doing this.
Providing Wildcard support would allow users to build extensions to
fulfill their own feature requests. It also would not necessarily
require much effort, since adding Wildcard support could be fairly
straightforward for a first-party that already has direct access to the
structured data in the page. (It would certainly be far simpler than
implementing a third-party site adapter.) There is also precedent for
first parties implementing an official client extension API: for several
years, Google maintained an official extension API in Gmail for
Greasemonkey scripts to use.

Google no longer maintains its own API, but third parties have continued
to maintain successful replacements---Gmail.js \autocite{talwar2019} is
an open source wrapper API with over 70 contributors and hundreds of
commits. This library demonstrates another benefit of designing for an
ecosystem: it's easier for many programmers to collectively maintain a
single adapter for a site and keep it updated in response to changes to
the site, rather than each programmer creating custom scraping logic for
their own individual extensions.

@@

@@

In the broadest sense, Wildcard is inspired by systems aiming to make
software into a dynamic medium where end users frequently create and
modify software to meet their own needs, rather than only consuming
applications built by programmers. These systems include Smalltalk
\autocite{kay1977}, Hypercard \autocite{hypercard2019} , Boxer
\autocite{disessa1986}, Webstrates \autocite{klokmose2015}, and
Dynamicland \autocite{victor}. (The project's name Wildcard comes from
the internal pre-release name for Hypercard, which doubly inspired our
work by promoting both software modification by end users and the ideas
behind the Web.)

While similar in broad goals, Wildcard employs a different solution
strategy. These projects generally require building software from
scratch in a new environment, whereas Wildcard aims to maximize the
malleability of software built with existing tools.

@@

Wildcard's solution approach is most similar to other tools that
identify structured data in a web page, and then use that structure to
support end user customization of the page.

@@

Thresher \autocite{hogue2005} enables users to create wrappers which map
unstructured website content to Semantic Web content. Like Wildcard and
Sifter, Thresher augments the page; once semantic content has been
identified, it creates context menus in the original website which allow
users to take actions based on that content. Wildcard and Thresher focus
on complementary parts of the customization process. Thresher aims to
enable end users to create content wrappers, but the actions available
on the structured data are determined in advance. Conversely, Wildcard
delegates wrapper creation to programmers, but gives end users more
flexibility to use the structured data in an open-ended way.

@@

\begin{itemize}
\tightlist
\item
  \emph{Reliability}: It can be difficult to know whether a command will
  consistently work over time. In addition to changes to the website,
  changes to the heuristics can also cause problems.
\item
  \emph{Discoverability}: it can be hard for users to find discover
  possible commands.
\end{itemize}

@@

Another difference between Wildcard and these tools is that there is no
way for end users to express imperative workflows with sequences of
actions in Wildcard. Chickenfoot and Coscripter have shown these types
of workflows to be useful in practice, but it's not clear how such
workflows could fit into Wildcard. Spreadsheets have a fundamentally
different computation model, and site adapters created by programmers
cannot easily account for every possible action a user might want to
perform in a page.

\hypertarget{spreadsheet-based-app-builders}{%
\subsection{Spreadsheet-based app
builders}\label{spreadsheet-based-app-builders}}

Many others have had the powerful realization that a spreadsheet can
serve as an backing data store and computation layer for an interactive
web application, enabling end users to create such applications more
easily. Research projects like Object Spreadsheets
\autocite{mccutchen2016}, Quilt \autocite{benson2014}, Gneiss
\autocite{chang2014}, and Marmite \autocite{wong2007}, as well as
commercial tools like Airtable \autocite{zotero-79} and Glide
\autocite{zotero-81} allow users to view data in a spreadsheet table,
compute over the data using formulas, and then connect the table to a
GUI. Because many users are already familiar with using spreadsheets,
this way of creating applications tends to be much easier than
traditional software methods; for example, in a user study of Quilt,
many users were able to create applications in under 10 minutes, even if
they expected it would take them many hours.

@@

An important design decision for tools in this space is how far to
deviate from traditional spreadsheets like Microsoft Excel or Google
Sheets. Quilt and Glide use existing spreadsheet software as a backend,
providing maximum familiarity for users, and even compatibility with
existing spreadsheets. Gneiss has its own spreadsheet implementation
with additional features useful for building GUIs. Marmite provides a
live data view that resembles a spreadsheet, but programming is actually
done using a separate data flow pane rather than spreadsheet formulas.
(Marmite's approach led to some confusion in a user study, because users
expected behavior more similar to spreadsheets \autocite{wong2007}.)
Airtable deviates the furthest: although the user interface resembles a
spreadsheet, the underlying structure is a relational database with
typed columns. Wildcard's table is most similar to Airtable; the
structure of a relational table is most appropriate for most data in
websites, and we have not yet found a need for arbitrary untyped cells.

@@

In this paper, we have presented \emph{spreadsheet-driven
customization}, a technique that enables end users to customize software
by augmenting an application's UI with a spreadsheet. We hope that this
technique contributes to making the Web into a more dynamic medium that
users can mold to their own needs.

There are still many open questions about spreadsheet-driven
customization which we hope to answer through targeted development and
usage of Wildcard.{\footnote{Up to date information about Wildcard is
  available on the project website
  (\url{https://geoffreylitt.com/wildcard}), which includes a signup for
  receiving updates and notifications about a public release. We also
  welcome hearing from readers who might be interested in being private
  beta testers or who have feedback on the project.}}

The most important question is: what are the limits of this
computational model? What types of useful customizations can it support
or not support? While initial demos suggest a variety of use cases, we
plan to develop more site adapters and demos to explore this question
further. We will start with privately testing the system with our own
needs, and then eventually deploy the tool publicly, once the API is
stable enough and can support a critical mass of sites and use cases. We
also plan to run usability studies to evaluate and improve the design of
the tool.

@@

Another open question is how easily site adapters can be created and
maintained for real websites, which often include complex markup and
change frequently. We plan to explore this question by creating adapters
for sites with different data structures and in different domains, and
perhaps running user tests with programmers. Possible future
improvements we've considered include developing automated heuristics to
assist with the adapter creation process, and developing new
abstractions that make it easier for programmers to efficiently create
new robust adapters.

>>>>

Web applications often don't meet the precise needs of their users.
Sometimes there is a browser extension available to patch an issue, and
if the user is a programmer they might be able to fix it themselves. But
for most people, the only recourse is to complain to the developers, or,
more likely, to simply give up. Back in 1977, in \emph{Personal Dynamic
Media} \autocite{kay1977}, Alan Kay envisioned personal computing as a
medium that let a user ``mold and channel its power to his own needs,''
but today, software behaves more like concrete than clay.

@@

We have implemented this technique in a prototype browser extension
called Wildcard and used it to build demos which suggest that this
paradigm can support useful customizations, ranging from sorting lists
of data to adding whole new features to applications (shown in
Section~\ref{sec:demos}).

@@

In addition to fetching data from other sources, Wildcard formulas can
also perform computations. In this example, the user would like to
augment the TodoMVC todo list app with a ``snooze'' feature, which will
temporarily hide a todo from the list until a certain date.{
Figure~\ref{fig:todomvc-demo} shows an overview of this customization.}

@@

In this section we've presented just a few use cases for
spreadsheet-driven customization, to suggest some possibilities of the
paradigm. We think the spreadsheet model is flexible enough to support a
wide range of other useful modifications while remaining familiar and
easy to use.

@@

  \CommentTok{// Function to extract DOM element:}
\NormalTok{  el}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \VariableTok{row}\OperatorTok{.}\FunctionTok{querySelector}\NormalTok{(}\StringTok{".\_1ebt2xej"}\NormalTok{)}\OperatorTok{,}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

Sometimes, more advanced scraping techniques are necessary to extract
data from the application. For example, we have prototyped a mechanism
for site adapters to observe AJAX requests made by the browser and
extract data directly from JSON responses. This mechanism was used to
implement the Airbnb Walkscore example, since latitude and longitude
aren't shown in the Airbnb UI, but they are available in AJAX responses.
This technique seems promising because AJAX responses already contain
data in a structured form, and web applications are increasingly loading
data using AJAX. Another technique we might consider is enabling site
adapters to scrape data across multiple pages for paginated lists of
results (as explored in \autocite{huynh2006}).

@@

First, the site adapter needs to support sending updates from the table
to the original page. We have made this fairly straightforward, because
most DOM manipulation is not performed directly by the site adapter.
Instead, Wildcard automatically mutates the DOM to reflect the
spreadsheet state, using the same specification that was used to extract
the data. The only exception to this is row actions (like favoriting an
Airbnb listing), which are implemented as Javascript functions that can
perform arbitrary behaviors like simulating clicks on buttons or
mutating the DOM.

Another challenge is triggering updates to the spreadsheet data in
response to UI changes that happen after initial page load. Site
adapters are responsible for recognizing these changes by observing the
DOM. So far, we have been able to use event listeners and the
MutationObserver API to successfully observe changes, but it may prove
challenging to observe changes on some sites only through the DOM.

@@

The idea of spreadsheet-driven customization is guided by three design
principles, inspired by prior work and our own experimentation. We think
these principles might also broadly inform the design of tools for end
user software customization.

@@

Computing does not need to be organized this way. For example, UNIX
offers a compelling alternative: many small single-purpose programs, all
of which manipulate a universal format of text streams. Users get high
leverage from mastering a text editor and other text manipulation tools,
since they can be applied to a huge variety of tasks. A user's preferred
text editor can even serve as an interactive input mechanism in shell
programs, e.g.~for editing git commit messages.

Spreadsheet-driven customization aims to bring some of this UNIX ethos
to the world of isolated Web applications by creating a consistent data
structure to represent the data inside many applications. In UNIX, the
universal format is a text stream; in Wildcard, it is a relational
table. Because Wildcard maps the data from all applications to the table
format, users can master the tools provided by Wildcard and apply them
to customize many different applications.

This idea relates to Beaudouin-Lafon and Mackay's notion of
\emph{polymorphic interface instruments} \autocite{beaudouin-lafon2000},
which are UI elements that can be used in different contexts (for
example, a color picker that can be used in many drawing applications).
It also relates to ideas of literacy in a medium. DiSessa notes that
textual literacy depends on the fact that writing can be adapted to many
different genres \autocite{disessa2000}: if people needed to relearn
reading and writing from scratch when switching from essays to emails,
the medium would lose its potency. Generic tools are especially
important for software customization, because a common barrier to
customizing software is not having enough time
\autocite{mackay1991}---it's more likely that people will customize
software frequently if they can reuse the same tools across many
applications.

This design principle leads to several challenges. First, any universal
abstraction has its limits, and may not naturally express the data in
every application. We plan to explore the limits of the table
abstraction further by trying to build adapters for more sites with
varied data formats. We expect that many types of data will fit easily
into tables: lists of search results, news articles, and messages can
all naturally be seen as relations. Documents (e.g.~Google Docs) or
graphs (e.g.~social network friend graphs) may prove more challenging to
map to the table abstraction.

@@

One of the most interesting properties of spreadsheets is that users
familiar with only a tiny sliver of their functionality (e.g., storing
tables of numbers and computing simple sums) can still use them in
valuable ways. This supports the user's natural motivation to continue
using the tool, and to eventually learn its more powerful features if
needed \autocite{nardi1991}. In contrast, many traditional programming
systems require an enormous upfront time investment before someone is
able to write a program that helps them achieve a useful task.

@@

Another aspect of providing a low floor is providing an ``in-place
toolchain'' \autocite{2019}: minimizing the effort of moving from using
to customizing by making customization tools available in the same
environment where the user is already using the software. This quality
is distinct from the level of technical skill needed to use the tool.
For example, setting up a workflow in an end user programming system
like \href{https://ifttt.com/}{IFTTT} does not require much programming
skill, but it does require leaving the user's normal software and
entering a separate environment; conversely, running a Javascript
snippet in the browser console requires programming skills, but can be
done immediately in the flow of using a website.

@@

Since we have only built a few site adapters and demos so far, it is
still too early to tell how high the ceiling is for the customizations
that can be achieved with Wildcard. But we think that with enough
built-in formulas (both standard spreadsheet formulas for computation,
and formulas for fetching data from other sources), the language could
support a wide variety of customizations. We plan to explore this aspect
further by trying to solve real problems and observing where limitations
emerge in practice (discussed more in Section~\ref{sec:conclusion}).

@@

Real-world spreadsheet usage is highly collaborative. Many users only
perform simple changes, while their coworkers help with writing complex
formulas or programming macros \autocite{nardi1990}. Inspired by this
collaborative dynamic, we aim for spreadsheet-driven customization to
bring together the abilities of different users.

@@

The first party developers of the original software can also play a role
in enabling customization. Although spreadsheet-driven customization
does not depend on cooperation from first-party website developers,
their participation in exposing structured data would eliminate the need
for third party site adapters. The first parties could themselves
benefit from such a system, since providing Wildcard support---a fairly
low-effort undertaking for a developer with direct access to the
structured data---would allow users to build extensions to fulfill some
of their own feature requests. There is also precedent for first parties
implementing an official client extension API: for several years, Google
maintained an official extension API in Gmail for Greasemonkey scripts
to use.

Google no longer maintains its own API, but third parties have continued
to maintain successful replacements---Gmail.js \autocite{talwar2019} is
an open source wrapper API with over 70 contributors and hundreds of
commits. This library demonstrates another benefit of designing for an
ecosystem: it's easier for many programmers to collectively maintain a
generic site adapter that can support many extensions, rather than each
programmer creating custom scraping logic for their own single-purpose
extension.

@@

Spreadsheet-driven customization relates to existing work in four areas:
malleable software, web customization, spreadsheet-driven applications,
and web scraping.

@@

In the broadest sense, Wildcard is inspired by systems aiming to make
software into a dynamic medium where end users frequently create and
modify software to meet their own needs, rather than only consuming
applications built by programmers. These systems include Smalltalk
\autocite{kay1977}, Hypercard \autocite{hypercard2019}, Boxer
\autocite{disessa1986}, Webstrates \autocite{klokmose2015}, and
Dynamicland \autocite{victor}. In fact, the name Wildcard comes from the
internal pre-release name for Hypercard, a project which doubly inspired
our work by promoting both software modification by end users and the
ideas behind the Web.

Although Wildcard shares broad goals with these projects, it employs a
different solution strategy. These other systems generally require
building software from scratch in a new environment. On the other hand,
Wildcard aims to maximize the malleability of software built with
existing tools.

@@

Wildcard's solution approach is most similar to other tools that
identify structured data in a web page and then use that structure to
support end user customization of the page.

@@

Thresher \autocite{hogue2005} enables users to create wrappers which map
unstructured website content to Semantic Web content. Like Wildcard and
Sifter, Thresher augments the page: once semantic content has been
identified, it creates context menus in the original website which allow
users to take actions based on that content. Wildcard and Thresher focus
on complementary parts of the customization process. Thresher aims to
enable end users to create content wrappers, but the actions available
on the structured data are determined in advance. Conversely, Wildcard
delegates wrapper creation to programmers, but gives end users more
flexibility to use the structured data in an open-ended way.

@@

\begin{itemize}
\tightlist
\item
  \emph{Reliability}: It can be difficult to know whether a command will
  consistently work over time. In addition to changes to the website,
  changes to the system's heuristics can also cause unexpected changes
  in behavior.
\item
  \emph{Discoverability}: it can be hard for users to find discover
  possible commands.
\end{itemize}

@@

Another difference between Wildcard and these tools is that there is no
way for users to express imperative workflows with sequences of actions
in Wildcard. Chickenfoot and Coscripter have shown these types of
workflows to be useful in practice so it might be worth considering how
to support them, but it's not clear how such workflows could fit into
Wildcard. Spreadsheets have a fundamentally different computation model,
and site adapters created by programmers cannot easily account for every
possible action a user might want to perform in a page.

\hypertarget{spreadsheet-driven-applications}{%
\subsection{Spreadsheet-driven
applications}\label{spreadsheet-driven-applications}}

Many researchers and companies have explored the idea of
\emph{spreadsheet-driven applications}: using spreadsheets as a backing
data store and computation layer for interactive web applications,
enabling end users to create such applications more easily. Research
projects like Object Spreadsheets \autocite{mccutchen2016}, Quilt
\autocite{benson2014}, Gneiss \autocite{chang2014}, and Marmite
\autocite{wong2007}, as well as commercial tools like Airtable
\autocite{zotero-79} and Glide \autocite{zotero-81} allow users to view
data in a spreadsheet table, compute over the data using formulas, and
then connect the table to a GUI. Because many users are already familiar
with using spreadsheets, this way of creating applications tends to be
much easier than traditional software methods; for example, in a user
study of Quilt, many users were able to create applications in under 10
minutes, even if they expected it would take them many hours.

@@

An important design decision for tools in this space is how far to
deviate from traditional spreadsheets like Microsoft Excel or Google
Sheets. Quilt and Glide use existing spreadsheet software as a backend,
providing maximum familiarity for users, and even compatibility with
existing spreadsheets. Gneiss has its own spreadsheet implementation
with additional features useful for building GUIs. Marmite provides a
live data view that resembles a spreadsheet, but programming is actually
done using a separate data flow pane rather than spreadsheet formulas.
Marmite's approach led to some confusion in a user study, because users
expected behavior more similar to spreadsheets \autocite{wong2007}.
Airtable deviates the furthest: although the user interface resembles a
spreadsheet, the underlying structure is a relational database with
typed columns. Wildcard's table is most similar to Airtable; the
structure of a relational table is most appropriate for most data in
websites, and we have not yet found a need for arbitrary untyped cells.

@@

In this paper, we have presented \emph{spreadsheet-driven
customization}, a technique that enables end users to customize software
by augmenting an application's UI with a spreadsheet. We still have many
open questions which we hope to answer through targeted development and
usage of Wildcard.{\footnote{Up to date information about Wildcard is
  available on the project website
  (\url{https://geoffreylitt.com/wildcard}), which includes a signup for
  receiving updates and notifications about a public release. We also
  welcome hearing from readers who might be interested in being private
  beta testers or who have feedback on the project.}}

The most important question is: what are the limits of this
computational model? What types of useful customizations can it support
or not support? While initial demos suggest a variety of use cases, we
plan to develop more site adapters and demos to explore this question
further. We will start by privately testing the system with our own
needs and then eventually deploy the tool publicly once it has a stable
API and can support a critical mass of sites and use cases. We also plan
to run usability studies to evaluate and improve the design of the tool.

@@

Another question is how easily site adapters can be created and
maintained for real websites, which often include complex markup and
change frequently. We plan to explore this question by creating adapters
for sites with different data structures and in different domains, and
perhaps running user tests with programmers. Possible future
improvements we've considered include developing automated heuristics to
assist with the adapter creation process and developing new abstractions
that make it easier for programmers to efficiently create robust
adapters.

We hope that spreadsheet-driven customization can help the Web reach its
full potential as a dynamic medium that users can mold to their own
needs.
